---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "85%",
  dpi = 300,
  fig.align = "center"
)
```


# dyadicdist <img src='man/figures/logo.svg' align="right" height="189" />

<!-- badges: start -->
[![CodeFactor](https://www.codefactor.io/repository/github/jvieroe/dyadicdist/badge)](https://www.codefactor.io/repository/github/jvieroe/dyadicdist)
[![Github All Releases](https://img.shields.io/github/downloads/jvieroe/dyadicdist/total.svg)]()
[![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![License: GPL (>= 3)](https://img.shields.io/badge/license-GPL (>= 3)-blue.svg)](https://cran.r-project.org/web/licenses/GPL (>= 3))
[![](https://img.shields.io/github/last-commit/jvieroe/dyadicdist.svg)](https://github.com/jvieroe/dyadicdist/commits/main)
<!-- [![Travis build status](https://travis-ci.org/jvieroe/dyadicdist.svg?branch=main)](https://travis-ci.org/jvieroe/dyadicdist) -->

Master branch status: [![R-CMD-check](https://github.com/jvieroe/dyadicdist/workflows/R-CMD-check/badge.svg)](https://github.com/jvieroe/dyadicdist/actions)
<!-- badges: end -->

# Introduction

The purpose of `dyadicdist` is to provide quick and easy calculation of dyadic distances between geo-referenced points. The main contribution of `dyadicdist` is that the output is stored as a long, dyadic `tibble` as opposed to a wide `matrix`.

This is still a development version. Please don't hesitate to let me know of any errors and/or deficiencies you might come across.

### Quick example
A simple example with illustrates the purpose of `ddist()` and the derived functions: `ddist_sf()`, `ddist_xy()`, and `ddist_xy_sf()`. 

```{r, warning=F, message=F}
library(tidyverse)
library(dyadicdist)

df <- tibble::tribble(
  ~city_name, ~idvar, ~latitude, ~longitude,
  "copenhagen", 5, 55.68, 12.58,
  "stockholm", 2, 59.33, 18.07,
  "oslo", 51, 59.91, 10.75
)

ddist(data = df,
      id = "idvar")

```


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
if(!require("devtools")) install.packages("devtools")
library(devtools)
devtools::install_github("jvieroe/dyadicdist")
```


# Usage: `ddist()`

Below, I describe some of the key features of `dyadicdist`. Let's use some data on the 100 largest US cities as a working example:

```{r, warning=F, message=F}
library(dyadicdist)
library(tidyverse)
library(magrittr)

cities <- dyadicdist::cities

```

`ddist()` has __two key inputs__. It requires a `data.frame` or `tibble` with specified latitude and longitude variables. Furthermore, it requires the specification of a unique id variable which can be either `numeric`, `integer`, `factor`, or `character`.

```{r, warning=F, message=F}
ddist(cities,
      id = "id")

```
As a default, latitude/longitude are specified as `"latitude"` and `"longitude"`, respectively, and don't need manual inputs. If necessary their variable names can be specified in the `ddist()` call:

```{r, warning=F, message=F}
cities %>%
  rename(lat = latitude,
         lon = longitude) %>% 
  ddist(.,
        id = "id",
        latitude = "lat",
        longitude = "lon") %>%
  head(5)

```

## Output specification
By default, `ddist()` returns the full list of dyadic distances between any points i and j, including j = i. 

In total, this amount to `nrow(data) * nrow(data)` dyads and includes by default:

* dyads between any observation and itself, i.e. dyads of type (i,i) (see example above)
* duplicated dyads, i.e. both (i,j) and (j,i)

Both of these inclusions are optional, however. 

* Sort out (i,i) dyads (the diagonal in a distance matrix) by specifying `diagonal = FALSE`
    + returns a `tibble` with `nrow(data) * (nrow(data)-1)` dyads
* Sort out duplicated dyads by specifying `duplicates = FALSE`
    + returns a `tibble` with `(nrow(data) * (nrow(data)-1)/2)+nrow(data)` dyads
* Sort out both by specifying `diagonal = FALSE` __and__ `duplicates = FALSE`
    + returns a `tibble` with `(nrow(data) * (nrow(data)-1)/2)` dyads


## CRS transformations
By default `ddist()` assumes unprojected coordinates in basic latitude/longitude format (EPSG code `4326`) when converting the raw data provided in the `data` argument to a spatial feature. This is consistent with the default when converting latitude/longitude data to spatial features in the `sf` package (see `sf::st_as_sf()`). You can apply a different CRS by providing a valid EPSG code of type `numeric` with the `crs` argument.

Additionally, `ddist()` allows you to transform the CRS before calculating dyadic distances using the `crs_transform` and `new_crs` arguments:

```{r, warning=F, message=F}
ddist(cities,
      id = "id",
      crs_transform = T,
      new_crs = 3359)

```

For a list of supported CRS transformations, see `rgdal::make_EPSG()`.

Note that the choice of CRS may impact your results considerably. For more information on choosing an appropriate CRS, see [here](https://www.earthdatascience.org/courses/earth-analytics/spatial-data-r/intro-to-coordinate-reference-systems/), [here](https://docs.qgis.org/3.4/en/docs/gentle_gis_introduction/coordinate_reference_systems.html), [here](https://www.nceas.ucsb.edu/sites/default/files/2020-04/OverviewCoordinateReferenceSystems.pdf), and [here](http://www.geo.hunter.cuny.edu/~jochen/gtech201/lectures/lec6concepts/map%20coordinate%20systems/how%20to%20choose%20a%20projection.htm)


## Spatial input data with `ddist_sf()`
To measure dyadic distances with an object of class `sf` use `ddist_sf()`:

```{r, warning=F, message=F}
library(sf)

cities %>%
  st_as_sf(.,
           coords = c("longitude", "latitude"),
           crs = 4326) %>%
  ddist_sf(.,
           id = "id")

```
With the exception of `crs`, `longitude`, and `latitude` (all of which are inherently provided in an object of class `sf`), `ddist_sf()` takes the same optional arguments as `ddist()`.

# Acknowledgements

* The R Core Team for developing and maintaining the language
* The authors of the amazing `sf` package. `sf` has greatly reduced barriers to entry for anyone working with spatial data in `R` and those who wish to do so 
    <!-- + Edzer Pebesma ([edzer](https://github.com/edzer)) -->
    <!-- + Roger Bivand ([rsbivand](https://github.com/rsbivand)) -->
    <!-- + Etienne Racine ([etiennebr](https://github.com/etiennebr)) -->
    <!-- + Michael Sumner ([mdsumner](https://github.com/mdsumner)) -->
    <!-- + Ian Cook ([ianmcook](https://github.com/ianmcook)) -->
    <!-- + Tim Keitt ([thk686](https://github.com/thk686)) -->
    <!-- + Robin Lovelace ([Robinlovelace](https://github.com/Robinlovelace)) -->
    <!-- + Hadley Wickham ([hadley](https://github.com/hadley)) -->
    <!-- + Jeroen Ooms ([jeroen](https://github.com/jeroen)) -->
    <!-- + Kirill Müller ([krlmlr](https://github.com/krlmlr)) -->
    <!-- + Thomas Lin Pedersen ([thomasp85](https://github.com/thomasp85)) -->
    <!-- + Dan Baston ([dbaston](https://github.com/dbaston)) -->
    <!-- + Dewey Dunnington ([paleolimbot](https://github.com/paleolimbot)) -->
* [LatLong.net](https://www.latlong.net/category/cities-236-15.html) for the `dyadicdist::cities` data
